---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# progressdatenbankderivate

<!-- badges: start -->
<!-- badges: end -->

The goal of progressdatenbankderivate is to extract data from an excel file 
containing an export from the database of the PROGRESS project and to compute 
several pneumonia scores based on this data.

## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("maciejrosolowski/progressdatenbankderivate")
```
## Example

Read the data exported from the database of the PROGRESS project.

```{r}
library(readxl)
library(data.table)
library(progressdatenbankderivate)
excel_fn <- paste0("/net/ifs2/san_projekte/projekte/", 
                   "PROGRESS/Datenmanagement/Data_freezes/", 
                   "20190320/PROGRESS-freeze_201903_01.xlsx")
DID_PROBAND <- readxl::read_excel(excel_fn, 'DID_PROBAND')
FRM_BAS <- readxl::read_excel(excel_fn, 'FRM_BAS')
FRM_BEF <- readxl::read_excel(excel_fn, 'FRM_BEF')
FRM_B24 <- readxl::read_excel(excel_fn, 'FRM_B24')
FRM_RR <- readxl::read_excel(excel_fn, 'FRM_RR', guess_max = 10e5)
FRM_O2A <- readxl::read_excel(excel_fn, 'FRM_O2A')
FRM_DIL_LABORWERTE <- readxl::read_excel(excel_fn, "FRM_DIL_LABORWERTE")
FRM_VIS <- readxl::read_excel(excel_fn, 'FRM_VIS')
DID_CLIN <- readxl::read_excel(excel_fn, 'DID_CLIN')
FRM_KAT <- readxl::read_excel(excel_fn, 'FRM_KAT')
FRM_O2P <- readxl::read_excel(excel_fn, 'FRM_O2P')
FRM_BEAT <- readxl::read_excel(excel_fn, 'FRM_BEAT', guess_max = 10e5)
DID_OXYGENIND_SINGLE <- readxl::read_excel(excel_fn, 'DID_OXYGENIND_SINGLE')
data.table::setDT(DID_PROBAND)
data.table::setDT(FRM_BAS)
data.table::setDT(FRM_BEF)
data.table::setDT(FRM_B24)
data.table::setDT(FRM_RR)
data.table::setDT(FRM_O2A)
data.table::setDT(FRM_DIL_LABORWERTE)
data.table::setDT(FRM_VIS)
data.table::setDT(DID_CLIN)
data.table::setDT(FRM_KAT)
data.table::setDT(FRM_O2P)
data.table::setDT(FRM_BEAT)
data.table::setDT(DID_OXYGENIND_SINGLE)
```

### PSI

This is an example which shows how to compute the Pneumonia Severity Index
(PSI):

```{r example}
# suppress warnings about no non-missing values while computing min or max
# by PATSTUID and EVENT.
suppressWarnings(
  psi_d0 <- psi.fct(DID_PROBAND, FRM_BAS, FRM_BEF, FRM_B24, FRM_RR, FRM_O2A, 
                    FRM_DIL_LABORWERTE,FRM_VIS, zp_fabian = "d0")$out[
                      , .(PATSTUID, EVENT, psi)]
)
psi_d0
```

### SIRS

```{r}
suppressWarnings(
  sirs_d0 <- sirs.fct(DID_PROBAND, FRM_BAS, FRM_BEF, FRM_B24, FRM_DIL_LABORWERTE,
                      DID_CLIN, FRM_RR, FRM_KAT, zp_fabian = "d0")$out[
                        , .(PATSTUID, EVENT, sirs = infec.septic.servsept,
                   schock = septischer.schock)]
)
sirs_d0
```

### quickSOFA

```{r}
suppressWarnings(
  qsofa <- quickSOFA(FRM_RR, FRM_B24, FRM_BEF, DID_CLIN,
                     zp_fabian = "auf_in_d0")$out
)
qsofa[, .(PATSTUID, EVENT, qSOFA)]
```

### Halm

```{r}
suppressWarnings(
  halm <- HalmScore(FRM_B24, FRM_BEF, FRM_RR, FRM_O2A, FRM_O2P, FRM_BEAT,
                    DID_CLIN, zp_fabian = "auf_in_d-1_in_d0")$out[
                      , .(PATSTUID, EVENT, halm)
                    ]
)
halm
```

### SCAP

```{r}
suppressWarnings(
  scap <- SCAP(FRM_B24, FRM_O2A, FRM_RR, FRM_BEF, FRM_DIL_LABORWERTE, DID_CLIN,
               DID_PROBAND, FRM_VIS, DID_OXYGENIND_SINGLE,
               zp_fabian = "auf_in_d-1_in_d0")$out[, .(PATSTUID, EVENT, SCAP)]
)
scap
```

### smartCOP

```{r}
suppressWarnings(
  smart_cop <- smartCOP(FRM_RR, FRM_BEF, FRM_VIS, FRM_DIL_LABORWERTE, FRM_B24,
                        DID_PROBAND, DID_CLIN, DID_OXYGENIND_SINGLE,
                        FRM_O2A, zp_fabian = "auf_in_d-1_in_d0")$out[
                          , .(PATSTUID, EVENT, smartCOP)]
)
smart_cop
```

### CRB, CRB65, CURB, CURB65

```{r}
suppressWarnings(
  curb65 <- curb65.fct(DID_PROBAND, FRM_BEF, FRM_B24, FRM_DIL_LABORWERTE, 
                       FRM_RR, FRM_BEAT = NULL, zp_fabian = "d0")$out[,
                         .(PATSTUID, EVENT, crb, crb65, curb, curb65)]
)
curb65
```


## How to compute the scores based on values of the input parameters

### PSI

```{r}
psi <- psi_simple(
  age = 68,
  verwirrt = 1,
  hfrq.max = 130,
  afrq.max = 35,
  sysbp.min = 80,
  temp.min = 34,
  temp.max = 35,
  tumor = 1,
  herz = 1,
  cerebro = 1,
  renal = 1,
  liver = 1,
  gender = 1,
  nurse.home = 1,
  art.ph.min = 7.2,
  bun = 10,
  snat = 120,
  gluk = 16,
  haemkrt = 0.2,
  apo2.min = 50,
  pleu_erg = 1)
psi
```

An example with vectors as arguments in the function call

```{r}
psi <- psi_simple(
  age = c(68, 61),
  verwirrt = c(1, 0),
  hfrq.max = c(130, 125),
  afrq.max = c(35, 26),
  sysbp.min = c(80, 98),
  temp.min = c(34, 36),
  temp.max = c(35, 37),
  tumor = c(1, 0),
  herz = c(1, 0),
  cerebro = c(1, 0),
  renal = c(1, 0),
  liver = c(1, 0),
  gender = c(1, 0),
  nurse.home = c(1, 0),
  art.ph.min = c(7.2, 7.5),
  bun = c(10, 11),
  snat = c(120, 134),
  gluk = c(16, 9),
  haemkrt = c(0.2, 0.25),
  apo2.min = c(50, 52),
  pleu_erg = c(1, 0))
psi
```

### SIRS

```{r}
sirs <- sirs_simple(
  temp.min = 34,
  temp.max = 35,
  hfrq.max = 104,
  afrq.max = 20,
  pco2 = 4.1,
  leuko_min = 7800,
  leuko_max = 7900,
  stkern.neutro = 395,
  smkern.neutro = 4424,
  verwirrt = 1,
  thrombo_min = 216000,
  thrombo.daybefore = 205000,
  oxi.ind = 43.81,
  chr.lunge = 1,
  diur = 3120,
  gewicht = 70,
  bemin = 3.4,
  sysbp.min = 95,
  map = 10.44,
  kate = FALSE)
sirs
```

### Halm

```{r}
halm <- HalmScore_simple(
  hfrq.max = 104,
  sysbp.min = 95,
  afrq.max = 19,
  o2p.min = 90,
  apo2.min = 50,
  bea = 1,
  sauerst = 0,
  temp.max = 38,
  verwirrt = 0,
  gcs = 15)
halm
```

